Comparing Generative Pre-trained Transformer Models to Physicians for Assigning Causes of Death from Verbal Autopsy: A Case Study of 6939 Deaths in Sierra Leone from 2019-2022

### 2024-07-06 manuscript/manuscript.tex
Working on discussion

\textbf{Results:} x.

@@

@@

@@

\begin{table}[h]
    \caption{Number of records with same GPT-3.5 outputs based on 10 repeated reruns of 100 records}\label{tab-reprod}%
    \begin{tabular}{@{}llll@{}}
    \toprule
    Times with Same GPT-3.5 Outputs & ICD-10 Records & CGHR-10 Records\\

@@

>>>>

\textbf{Results:} Based on PCCC, GPT-4 had the best performance overall (0.61), followed by GPT-3.5 (0.58), InSilicoVA (0.44), and InterVA-5 (0.43). GPT-4 also had the best performance for the adult (0.65), child (0.57), and neonatal (0.62) age groups, while performance decreased as adults aged (0.72-0.59) and increased as children and neonates aged (0.5-0.7). 7 of 19, 4 of 10, and 1 of 7 adult, child, and neonatal CODs respectively had models with $\geq$0.8 PCCC. For most CODs, GPT models had one of the highest performances with the exception of InSilicoVA for adult tuberculosis, child pneumonia and neonatal infections. At the population level, all models had CSMF accuracy between 0.7-0.79.

@@

This section discusses research opportunities to improve GPT models for assigning CODs. Section \ref{discuss-opp-icd10} suggests comparing GPT models with physicians using ICD-10 codes. Section \ref{discuss-opp-peng} discusses the potential to improve GPT models with prompt engineering. Finally, an exploration of misclassified cases with domain knowledge is discussed in Section \ref{discuss-opp-misclass}.

@@

\subsubsection{Explore Misclassifications}\label{discuss-opp-misclass}

@@

\begin{table}[h]
    \caption{Number of records with same GPT-3.5 outputs based on 10 repeated reruns of 100 records}\label{tab-reprod}%
    \begin{tabular}{@{}lll@{}}
    \toprule
    Times with Same GPT-3.5 Outputs & ICD-10 Records & CGHR-10 Records\\

@@

\clearpage
\section{Exploration of Neonatal Infections}\label{app-misclass}

An exploration of neonatal infections (n=99, 21\% of 477 records) was done to understand the low performance of GPT models (0.23 PCCC) for neonatal infections, and high performance of InSilicoVA (0.87 PCCC). In Table \ref{tab-misclass}, about half the records were assigned correctly, and a majority (n=33, 33\%) of the other records were misclassified as other, while prematurity and low birthweight, birth asphyxia \& birth trauma, and ill-defined make up the rest. On closer inspection of the 49 records with misclassified assignments, the ICD-10 code R50 was assigned in 20 records. R50 falls under unspecified infections in the adult CGHR-10 category, but in the other category for neonates. B50 was assigned in 4 records, falling under malaria, but a similar B54 falls under neonatal infections. P81 was assigned in 3 records, referring to fever of unknown origin, which falls under other, and P07 was assigned in 7 records, falling under prematurity and low birthweight.

In most misclassified records, there is mention of infections, but the misclassifications occur due to the finer details of the ICD-10 code classifications, the categorization decisions of the CGHR-10 codes, and missing information from the questionnaire. For R50 misclassifications, GPT may have confused descriptions across adult and neonatal age groups. Using the same definition of R50, but in the context of neonates, may result in codes closer to neonatal infections (e.g. B54). For B50 misclassifications, the similar B54 was categorized in CGHR-10 as neonatal infections, but B50 was categorized as other. P81 refers to fever of unknown origin, which may be difficult to differentiate between infection and other causes without information from the questionnaire. P07 refers to prematurity and low birthweight, where GPT initially assigned P07 as the age of the neonate was mentioned first, but later mentions infections as an alternative following the order of information in the narratives. Thus, it may be possible to improve the performance GPT models using better prompts based on the context of VA manuals and CGHR-10 codes, and by also including questionnaire information in the prompts.

\begin{table}[h]
    \caption{CGHR-10 cause of death assignment from GPT-4 for records coded as neonatal infections by physicians}\label{tab-misclass}%
    \begin{tabular}{@{}ll@{}}
    \toprule
    GPT-4 Assigned Cause of Death (CGHR-10) & Records\\

    Neonatal infections & 50 (51\%)\\
    Other & 33 (33\%)\\
    Prematurity and low birthweight & 9 (9\%)\\
    Birth asphyxia \& birth trauma & 5 (6\%)\\
    Ill-defined & 2 (2\%) \\

    Total & 99 (100\%)\\

    \botrule
    \end{tabular}
\end{table}
